### 2019-04-15 presentation/pres.tex
add figures and results

\documentclass[slidestop,compress,red,mathserif]{beamer}

\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{amsmath,amssymb}
\usepackage{comment}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{marvosym}
\usepackage{natbib}
\usepackage{ulem}
\usepackage{wasysym}

\setbeamersize{text margin left=6mm, text margin right=2mm}

\mode<presentation>
{
%\usetheme{Warsaw}
%\usetheme{Rochester}
%\usetheme{Madrid}
%\usetheme{Pittsburgh}
%\usetheme{Antibes}
%\usetheme{Montpellier}
%\usetheme{Berkeley}
%\usetheme{PaloAlto}
%\usetheme{Goettingen}
%\usetheme{Marburg}
%\usetheme{Hannover}
%\usetheme{Berlin}
\usetheme{Ilmenau}
%\usetheme{Dresden}
%\usetheme{Darmstadt}
%\usetheme{Frankfurt}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Copenhagen}
%\usetheme{Malmoe}
\setbeamercovered{transparent}
}

% Choose color scheme
%
\usecolortheme{default}
%\usecolortheme{sidebartab}
%\usecolortheme{albatross}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{seagull}
%\usecolortheme{lily}
%\usecolortheme{orchid}

\beamertemplateballitem
%\setbeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\defbeamertemplate*{footline}{shadow theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=2.\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fil,rightskip=.3cm]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertframenumber\,/\,\inserttotalframenumber\hfill%\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    \usebeamerfont{title in head/foot}%\insertshorttitle%
  \end{beamercolorbox}}%
  \vskip0pt%
}

% number sets
\def\N{\hbox{I \kern-.5em N}}
\def\R{\hbox{I \kern-.5em R}}
\def\Q{\hbox{\hspace*{.25em}{\sf I} \kern-.83em Q}}
\def\T{\rm T}  %%%%  \rm pour roman (ecriture normale)
\def\Z{\rm Z}
\def\ZZ{\hbox{\sf  Z \kern-.8em Z}}
\def \Zplus        {\ZZ^{+}}
\def \bigm {\textsc{penal}}
\def \dst {\textsc{dst}}
\def \src {\textsc{src}}
\def \zlp {z^{\star}_{\textsc{lp}}}
\def \zilp {{\tilde z}_{\textsc{ilp}}}
\def \l    {\textnormal{\textsc{l}}}
\def \Gp   {G_{\textsc{p}}}
\def \Gl   {G_{\l}}
\def \Vp   {V_{\textsc{p}}}
\def \Vl   {V_{\l}}
\def \Ep   {E_{\textsc{p}}}
\def \El   {E_{\l}}
\def \cost {\textsc{cost}}
\def \srlg {\textsc{srlg}}

% number sets
\def \best {\textsc{best}}
\def  \N            {\hbox{I \kern-.5em N}}
\def \RR {\mathbb{R}}
\def \Rplus {\RR^+}
\def  \Z            {\rm Z}
\def  \Zplus        {\Z^+}
\def  \Zplusn       {\Zplus_n}
\def  \ZZ           {\hbox{\sf  Z \kern-.8em Z}}
\def \Z {\mathbb{Z}}

\logo{%
    \includegraphics[width=2.cm,keepaspectratio]{Figures/ConULogo_CMYK.eps}~%
}

\title[My Title]{Accident prediction in time and space}
\author[My~Name]{
\textbf{Timothée Guédon, Antoine Hébert}}
\institute[My Institute/company]
{Concordia University \\ Computer Science and Software Engineering Department}
\date[]{April 15, 2019}

\begin{document}

\begin{frame} % Cover slide
\titlepage
\end{frame}

% layout of the presentation
\AtBeginSubsection[]
{
  \begin{frame}<beamer>
    \frametitle{Layout}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

%-------------------------------------------------
\section{Introduction}
%-------------------------------------------------

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Road accident prediction}

\begin{scriptsize}

\begin{itemize}
\item[] \textbf{Context}:
  \begin{itemize}
    \item Road accidents result in millions of people deaths and injuries every year.
		\item Human and economic impact on society.
    \item Master’s thesis of Antoine Hébert in a different context.
  \end{itemize}

\item[] \textbf{Our goals}:
  \begin{itemize}
    \item Dataset assembling.
    \item Predict the risk of road accident at a given time on a given road.
  	\item Finding key insights in accident prediction.
  	\item Identifying issues related to class imbalance and geo-spatial data.
  \end{itemize}
\end{itemize}

\end{scriptsize}

\end{frame}

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Datasets}

\begin{footnotesize}
We used open datasets from Open Canada.

\begin{itemize}
\item[] \textbf{National Collision Database (NCDB)}:
  \begin{itemize}
  	\item All vehicle collisions reported by the police from 2012 to 2017 in Montreal.
  	\item Date and time, severity, number of death and injury, number of vehicles, localization, speed limit etc.
  \end{itemize}
\item[] \textbf{National Road Network}:
  \begin{itemize}
    \item Shape of the road, how roads are
connected together, the mail addresses on the street, etc.
  \end{itemize}
\item[] \textbf{Climate data - Hourly Data Report}:
  \begin{itemize}
    \item The dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc. at different weather stations in Montreal.
  \end{itemize}
\end{itemize}

\end{footnotesize}

\end{frame}

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Tree-based algorithms}
We chose to focus on tree-based algorithms because it has proven its efficiency on accident prediction.
\begin{itemize}
\item[] \textbf{Related work on accident prediction}:
  \begin{itemize}
    \item Extensive use of decision trees in several form. (\cite{Theofilatos2017}, \cite{Abellan2013}, \cite{Lin2015}, \cite{Chang2005})
    \item Recently, deep learning algorithms (LSTM and CNN architectures for example) (\cite{Yuan2018}, \cite{Chen2016})
  \end{itemize}
\item[] \textbf{Chosen algorithms}:
  \begin{itemize}
  	\item Random Forest
    \item Balanced Random Forest
    \item Gradient Boosted Trees (GBT) - XGBoost implementation
  \end{itemize}
\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Methods}
%-------------------------------------------------

\begin{frame}
	\frametitle{Generating the examples}
  \begin{itemize}
  \item[] \textbf{Positive samples}:
    \begin{itemize}
      \item Identification of road for each accident: Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
		  \item Extrapolation of missing points on the roads for more precise road identification
    \end{itemize}
  \item[] \textbf{Negative samples}:
    \begin{itemize}
      \item Every time an accident did not happen is considered a negative sample.
  		\item We extracted a subsample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
  		\item Generation of 5 million negative samples for about 200 000 positive samples.
    \end{itemize}

	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Some feature engineering}
	\begin{itemize}
		\item Transformation of the dates and times into a cyclic format
    \item Interpolation of weather statistics at a given point by averaging data collected from several stations
    \item Road features: street type and length
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Area under PR curve}
	\begin{itemize}
    \item ROC Curve : false positive rate VS true positive rate
    \begin{equation*}
			FPR = \frac{FP}{FP+TN},  TPR = \frac{TP}{TP+FN}
		\end{equation*}
		\item PR Curve : true positive rate VS precision
    \begin{equation*}
			TPR = \frac{TP}{TP+FN},  Precision = \frac{TP}{TP+FP}
		\end{equation*}
    \item Data imbalance imply decrease of TP and increase of FN (none of them appears in the FPR)
    \item Therefore, area under PR exhibits larger differences than area under ROC (\cite{Davis2006})
	\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Balanced Random Forests}
%-------------------------------------------------

\begin{frame}
\frametitle{Remainder: Random Forests}
\begin{itemize}
  \item[] \textbf{Remainder}
    \begin{itemize}
    	\item Random Forests (\cite{Breiman2001}) are combinations of fully-grown decision trees.
    	\item Each tree is trained on a random sample of the data and a random sample of the features.
      \item The final prediction is given by averaging the votes of the trees.
    \end{itemize}

  \item[] \textbf{Weighted Random Forests (WRF)}:
    \begin{itemize}
      \item A solution to imbalance (\cite{Chen2004}).
      \item Add weights to the classes when growing the tree (in the Gini coefficient).
      \item Add weights to the classes at the voting time.
    \end{itemize}

\end{itemize}
~~~~~~~~~~~~~~~
\end{frame}

\begin{frame}
\frametitle{Solution to imbalance}
\begin{itemize}
  \item[] \textbf{Balanced Random Forests (BRF)}:
    \begin{itemize}
      \item Another solution to imbalance (\cite{Chen2004}).
      \item Combine undersampling and ensemble method to migitate the loss of information.
      \item For each tree : Randomly draw samples of same size from majority (with replacement) and minority class.
      \item The rest of the algorithm is identical to Random Forests
    \end{itemize}

  \item[] \textbf{BRF versus WRF}:
    \begin{itemize}
      \item No clear winner in terms of prediction power ! But...
      \item WRF - More vulnerable to noise on minority class (\cite{Chen2004})
      \item BRF - More computationally efficient because use less data (subsampling) (\cite{Chen2004})
      \item BRF - Easier to implement in Spark (see next slide)
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Implementation of BRF in Spark}
\begin{itemize}
  \item[] \textbf{Existing implementation of RF in Spark}
    \begin{itemize}
    	\item To manage the subsampling of the dataset, Spark's scala implementation use the Poisson distribution.
      \item The Poisson distribution is used to compute the probability of an event to occur at a given time.
    	\item To know how many times a given sample will be in the subsample of a given tree (can be 0 times), we use "distribution sampling" of 1 element. (Distribution sampling: Get a random sample following a given distribution)
    \end{itemize}
  \item[] \textbf{Our modifications}:
    \begin{itemize}
      \item Instead of creating one Poisson distribution (with the "p" parameter and a seed) we create two Poisson distributions:
    	\item[]
      \begin{itemize}
      	\item One for the distribution of the "positive" samples,
        \item Another for the distribution of the "negative" samples.
      \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Experimenting with our BRF implementation}
\begin{itemize}
  \item Area Under PR : 98,1\% for RF VS 98,7\% for BRF
  \item F1 score : 99.7\% for RF VS 98.8\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)
  \item With BRF, we detect more accidents but it increases the false positive rate.

\end{itemize}

\centering
\includegraphics[height=3.0cm, keepaspectratio]{Figures/RF.png}
\includegraphics[height=3.0cm, keepaspectratio]{Figures/BRF.png}

\end{frame}

%-------------------------------------------------
\section{Big Data related issues}
%-------------------------------------------------

\begin{frame}
	\frametitle{Dataset generation}
	\begin{itemize}
		\item Using cluster to increase the dataset generation's speed.
    \item Tricks to increase the generation speed and reduce the memory consumption.
	\end{itemize}
\centering
\includegraphics[height=5.0cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
\end{frame}

\begin{frame}
	\frametitle{Spark VS Dask}
	\begin{itemize}
    \item[] \textbf{Advantages of using Dask over Spark}
      \begin{itemize}
      	\item Mimic the pandas dataframe API
        \item Lighter than Spark
      	\item Designed to be more flexible in terms of applications and algorithms
      \end{itemize}
    \item[] \textbf{Why we switched for Spark}
      \begin{itemize}
        \item We found ourselves writing a lot more code using Dask than using Spark
        \item We found Spark easier to use for processing the dataset in batch
        \item Spark's map-reduce and shuffling operations are very handy
      \end{itemize}
	\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Results}
%-------------------------------------------------

\begin{frame}
	\frametitle{Results}
	\begin{itemize}
		\item Random forest: 15.3\% area under PR curve
		\item Balanced random forest: 17.2\% area under PR curve
	\end{itemize}
\centering
\includegraphics[height=4cm, keepaspectratio]{Figures/rf_results.png}
\includegraphics[height=4cm, keepaspectratio]{Figures/nohyperparam_brf.png}
\end{frame}

\begin{frame}
	\frametitle{Key features identified}
	\begin{itemize}
		\item The most important features seem to be the time and location of the vehicle, followed by some road's features like the type of road (highway for example) and the street length.
    \item Finally, come some weather features like the visibility, the wind and the humidity.
	\end{itemize}
\centering
\includegraphics[height=4cm, keepaspectratio]{Figures/features.png}
\end{frame}

%-------------------------------------------------
\section{Conclusion}
%-------------------------------------------------

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
    \item We found some key features that can explain the occurence of accidents.
    \item We identified some solution to deal with class imbalance and big data-related issues.
    \item We implemented BRF in Spark.
    \item Future work would include tuning the XGBoost classifier, adding some datasets like a road works dataset.
    \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal
	\end{itemize}
\end{frame}

% -------------------------------------------------------------------------------------------------------------------------
% -------------------------------------------------------------------------------------------------------------------------
%% -------------------------------------------------------------------------------------------------------------------------

\begin{frame}[allowframebreaks]
\frametitle{References}

{\scriptsize

\bibliographystyle{plainnat}
\bibliography{Biblio}

}

\end{frame}

\end{document}